\center \fontsize{18.8}{20} \selectfont Towards Globally Color-Dual Multi-Loop Integrands

### 2023-09-22 notes/2loopNLSM.tex
Done w readthrough

\usepackage[notcite]{showkeys}
\usepackage{tikz-feynman} 
\usepackage{xcolor}
\hypersetup{
    colorlinks,
    linkcolor={nhpRed},
    citecolor={nhpBlue},
    urlcolor={nhpBlue}
}

@@

As noted above, we can select out can manifestly cubic
sector of semi-abelian Yang-Mills from the full theory of \cref{eq:ym-lag} by selecting only $(\varepsilon^+\varepsilon^+)$ in light-cone gauge (i.e. SDYM) and one-minus at tree-level or by plugging in the on-shell states of $J$-theory \cite{Cheung:2020djz,Cheung:2021zvb} and
$YZ$-theory \cite{Cheung:2016prv}. Indeed, semi-YM is just the following blocks
in the expansion of Yang-Mills given in Ref.~\cite{Pavao:2022kog}
\begin{equation}
A(A_1,...,\bar{A}_i,...,A_n)_{\text{tree}} = \sum_{i\neq j} (\varepsilon_i \varepsilon_j) \Delta^{(ij)}_{\text{tree}} \sim \sum_{i\neq j} (\varepsilon_i \varepsilon_j)  \sum_a \,[(\varepsilon k)^{n-2} (kk)^{3-n}]_a\,,
\end{equation}
and similarly so at one-loop
\begin{equation}
A(A_1,...,A_n)_{\text{1-loop}} = \Delta^{(\varnothing)}_{\text{1-loop}} \sim \sum_a\,[ (\varepsilon k)^{n} (kk)^{-n}]_a\, .
\end{equation}
where $[\,\cdots]_a$ are terms with appropriate powers of $(\varepsilon k)$ and $(kk)$. As we describe in detail in \cref{sec:CKLagrangians}, to recover NLSM at one-loop we need to extract the $D$-dependent part
that corresponds to an internal $\bar{Y}Y$-loop from extra-dimensional
scalars
\begin{eBox}
\begin{equation}
A^{\text{NLSM}}_{\text{1-loop}} \equiv \partial_D \Delta^{(\varnothing)}_{\text{1-loop}}\big|_{\varepsilon\rightarrow k}\, .
\end{equation}
\end{eBox}
Why must we take the derivative with respect to $D$? After all,
$J$-theory and $YZ$-theory both have well defined propagators for
internal $\bar{J}J$ and $\bar{Z}Z$ propagators. However, as we discuss
in \cref{sec:CKLagrangians}, all the $J$ and $Z$ states must be on-shell in order to
produce NLSM amplitudes. Thus, the unitarity cuts of $J$ theory at
one-loop will not produce NLSM amplitudes. However, the internal
$YY$-loop is a valid forward limit for producing NLSM amplitudes,
since as constructed $YZ$-theory matches to NLSM for off-shell
Y-particles. We now provide explicit expressions for these one-loop amplitudes in the next section.

\subsection{One-loop color-dual integrands}\label{oneLoopCK}
Our first application of this theory for color-dual construction at
loop level is for self-dual Yang-Mills (SDYM). As we have done throughout the text, we will set all coupling constants to unity. At tree-level, the
off-shell cubic vertices can be written in terms of light cone
coordinates \cite{Monteiro2011pc}:
\begin{equation}
X(p,k) = p_u k_w-p_w k_u
\end{equation}
At loop-level, this construction needs to be analytically continued to
general dimension in order to apply dimensional regularization at
one-loop. This can be done using the cubic semi-YM vertices of the
previous section:
\begin{equation} \label{DdimSDYM}
\cubic{hgrey0}{fermion2}{}{fermion2}{}{fermion2}{}= \mathcal{X}(k_1,k_2) = (\varepsilon_1\bar{\varepsilon}_3)(\varepsilon_2 k_3) -  (\varepsilon_2\bar{\varepsilon}_3)(\varepsilon_1 k_3)
\end{equation}
Plugging in on-shell all-plus helicity states in light-cone gauge will
yield precisely the 4D SDYM vertex, up to an unphysical phase:
  \begin{equation}
  \mathcal{X}(k^+_1,k^+_2) = \frac{\langle 12\rangle^3}{ \langle23\rangle \langle 31\rangle} \sim \langle 12\rangle = k_{1,u} k_{2,w}-k_{1,w} k_{2,u}
 \end{equation}
where $\mathcal{X}(k^+_1,k^+_2) = A^{\text{YM}}(1^+,2^+,3^-)$ in light cone gauge, and in the second step applied momentum conservation to redefinition of the spinor bracket, $\langle 12\rangle \rightarrow X(p,k)$ defined in \cref{Conventions}. Of course, the form of \cref{DdimSDYM} has the advantage of
permitting a $D$-dimensional construction of the one-loop
integrand. Below we provide a four-point example:
\begin{equation}
\simplebox = \langle \mathcal{X}(k_1,\ell_1)\mathcal{X}(k_2,\ell_2)\mathcal{X}(k_3,\ell_3)\mathcal{X}(k_4,\ell_4)\rangle
\end{equation}
where $\ell_i = \ell-(k_1+k_2+\cdots+ k_i)$ and the bracket $\langle \,\cdots\rangle$ indicates that we have applied the gauge fixed state projector, $\sum {\varepsilon^\mu_{(+\ell_i)}\varepsilon^\nu_{(-\ell_i)}}=\eta^{\mu\nu}$ on all internal polarizations.
In general, the $n$-gon diagram would be,
\begin{equation}\label{SDYMnGon}
N^{\text{SDYM}}_{n\text{-gon}}= \nGonVector = \langle \mathcal{X}(k_1,\ell_1)\mathcal{X}(k_2,\ell_2)\cdots \mathcal{X}(k_n,\ell_n) \rangle
\end{equation}
Due to the state-sum of internal loop factors,
$\sum {\varepsilon_{(+\ell)}\varepsilon_{(-\ell)}}\sim D$, the integrand above
depends explicitly on the spacetime dimension, $D$. By taking a
derivative\footnote{Another way to understand this derivative is that
  it selects the large $D$ behavior of the integrand.}, we can recover
the integrand needed for all-plus one-loop amplitudes with an internal
scalar:
\begin{equation}
\partial_D N^{\text{SDYM}}_{n\text{-gon}}=\nGonScalar  = 2^n (\ell_1 \varepsilon_1)(\ell_2 \varepsilon_2)\cdots (\ell_n \varepsilon_n)
\end{equation}
This integrand numerator with internal scalar loop is precisely what
one would obtain from the Feynman rules of $YZ$-theory, absent the $\bar{Z}Z$
internal vector loop. For more background, we refer the reader to
\cref{sec:CKLagrangians}. When plugging on the on-shell states of $YZ$-theory, we thus obtain the following expression for the NLSM one-loop $n$-gon numerator:
\begin{equation}\label{NLSMnGon}
N^{\text{NLSM}}_{n\text{-gon}}=\left[\partial_D  N^{\text{SDYM}}_{n\text{-gon}}\right]^{\epsilon \rightarrow k}=\nGonScalar =   [12][23]\cdots [n1]
\end{equation}
Where we have define the antisymmetric kinematic variable,
$[ij] = \ell_i^2 - \ell_j^2$, since when plugging on longitudinal
modes, the $n$-gon can be re-expressed completely in terms of inverse
propagators, $2(k_i \cdot \ell_i) = \ell_{i}^2-\ell^2_{i-1}$. We have
verified through 10-point one-loop that this $n$-gon expression is a
valid color-dual representation for NLSM.

Before proceeding, we note that the above definition does give rise to ``pathological" bubble-on-external-leg (BEL) diagrams, discussed previously in \cref{sec:bootstrap}. However, one can show that these diagrams integrate to zero for spacetime dimension, $D>2$, and thus can be disregarded as unphysical. For the interested reader, in \cref{BELreg} we provide a detailed overview of this dimensional regularization of the BEL diagram.
\subsection{Two-loop obstruction}\label{2LoopObstruction}
At two-loop, introducing terms that conspire with four-point contacts
is unavoidable. At one-loop, we were able to avoid internal
contractions of $\bar{A}_\mu A^\mu$ by selecting appropriate external
states. However, at two-loop when choosing all external $A_\mu$
states, the amplitude vanishes in semi-YM theory:
\begin{equation}
\mathcal{A}^{\text{semi-YM}}_{\text{2-loop}}(A_\mu,A_\mu,A_\mu,A_\mu)=0
\end{equation}
In order to produce non-vanishing interactions, we would need to
reintroduce $D$-dimensional vertices from the diagonal group of
$G_L\times G_R$ that have the wrong weight. Doing so will necessarily permit internal
contractions of $\bar{A}_\mu A^\mu$,
\begin{equation}\label{2loopTension}
\mathcal{A}^{\text{YM}}_{\text{2-loop}}(A_\mu,A_\mu,A_\mu,A_\mu)=\doubleBoxObstructionA +\doubleBoxObstructionB+\cdots 
\end{equation}
where we used white dots to indicate interaction vertices of
weight\footnote{We choose the convention that the abelianized gauge
  fields carry weight $\mathcal{W}[\bar{A}] = -1$, and non-abelian
  vectors are $\mathcal{W}[\bar{A}] = +1$.}
$\mathcal{W}[\bar{A}\bar{A} A] = -1$, rather than the isolated semi-YM
vertex which always carries weight $\mathcal{W}[\bar{A}A A] =
+1$. This immediately runs into the difficulty of introducing the
four-point contact needed for color-kinematics to be satisfied on all
internal edges. By a simple weight counting argument, one can see that
including this wrong sign interactions is unavoidable at two-loop and
higher.

Thus, to construct two-loop numerators prescriptively, as we have done at one-loop, would require knowledge of
the full kinematic algebra for Yang-Mills off-shell. Since this is
presently unavailable, to approach the problem of two-loop
construction, we instead employ an ansatz based approach.

\section{Two-loop four-point bootstrap}
\label{2loopBoot}
As we have just seen, it is possible to coerce tree numerators into
1-loop numerators at any multiplicity for pions (and related
theories), but that these methods cannot be reapplied to generate
higher-loop numerators.  For pions in particular, the no-go statement
is only for \emph{one particular representation of the theory} so does
not completely preclude the existence of a two-loop color-dual
representation.  We thus turn to the color-dual bootstrap method
described in \cref{sec:bootstrap} to construct a color-dual
representation of two-loop four-point NLSM.  Because of the similarity
of the problem setup, we will also use the opportunity to revisit the
construction of a color-dual representation of pure YM, extending the
search space beyond what was covered in Ref.~\cite{Bern:2015ooa}.

Both representations share the same cubic graph basis, and thus their
defining Jacobi relations lead to the same numerator basis.
Ignoring tadpoles and BEL graphs, there are 14 cubic four-point two-loop
graphs (see \cref{fig:MaxCuts}) which are related to each other via 21
Jacobi relations.  As mentioned in \cref{sec:bootstrap},
color-kinematics duality ensures that the numerator of every graph can
be expressed in terms of a basis of the double box and penta-triangle,
\begin{equation}
\label{eq:jacobi-basis}
\NLSMBasisDoubleBox \text{ and } \NLSMBasisPentaTriangle .
\end{equation}
However, the actual numerator dressings and physical properties
differ, so we discuss them separately below.

@@

For a scalar theory like NLSM the numerator only depends on dot
products of momenta.  The basis of momentum invariants for both of the
graphs can be taken to be
\jm{This looks beautiful!  Thank you to whoever formatted this.}
\begin{equation}
\label{eq:scalar-basis}
N^{\text{NLSM}}\left[
  \scalebox{0.75}{\NLSMBasisDoubleBox} \right]
\text{ and }
N^{\text{NLSM}}\left[\scalebox{0.75}{\NLSMBasisPentaTriangle} \right]\in
\left\{\begin{array}{cccc}
    k_{13},& k_{23},& k_{15},& k_{16}, \\
    k_{25},& k_{26},& k_{35},& k_{36}, \\
    k_{55},& k_{56},& k_{66}
  \end{array}
  \right\},
\end{equation}
where $k_{ij}$ means $k_i \cdot k_j$, $k_5=\ell_1$, and $k_6 = \ell_2$.
%%%%
% Basis for both diagrams
% {d[p[1], p[3]], d[p[1], p[6]], d[p[1], p[7]], d[p[2], p[3]],  d[p[2], p[6]], d[p[2], p[7]], d[p[3], p[6]], d[p[3], p[7]],  d[p[6], p[6]], d[p[6], p[7]], d[p[7], p[7]]}
%%%%
Every vertex in NLSM scales as $k^2$ so the numerator of each diagram
scales as $k^{12}$.  At this mass dimension there are 8,008
independent dot products of momenta.  With two basis graphs there is a
total of 16,016 ansatz parameters.\footnote{\texttt{Mathematica} is
  not suited to solving large sets of equations with sixteen thousand
  parameters, so a different method is desirable.  One possibility is
  to use a sprase solver, as often employed for integral reduction.
  However, these solvers typically assume that the matrix has a very
  low density.  While the equations in YM naturally separate into
  different ``helicity'' sectors (or $e_i \cdot e_j$ and
  $e_i \cdot k_j$ sectors in $D$ dimensions) which limits the
  cross-talk between equations and thus the effective density, there
  is no such separation of sectors in a scalar theory.  Thus the NLSM
  constraint equations were solved using a custom solver, working name
  \texttt{FiniteFieldSolve}, designed to exactly solve large linear
  systems of arbitrary density.  \texttt{FiniteFieldSolve} will be
  released publicly shortly.} Imposing the boundary Jacobi relations
eliminates all but 4,473 of the original free parameters.  Further
imposing graph symmetries reduces the number of free parameters to
1,243.

The final essential constraint is unitarity.  The two physical cuts
that need to be enforced, shown in \cref{fig:emu}, only involve
four-point vertices.  Notably this means that even at two loops,
unitarity is still only probing the on-shell four-point NLSM amplitude.
Unitarity cuts besides those shown in \cref{fig:emu} are either
pathological or vanish because they involve a three-point point on-shell
amplitude. In order to manifest the $\mathbb{Z}_2$
symmetry of NLSM at the integrand level, we require that all
non-pathological cuts containing a three-point vertex (such as the maximal
cut) vanishes.  Imposing all relevant cuts leaves 765 free parameters
representing the generalized gauge freedom in the answer.

@@

At this point color-kinematics duality has been satisfied, so the
double copy will proceed without issue.  However, the plethora of
generalized gauge freedom parameters leaves the option for enforcing
additional aesthetic constraints.  All remaining generalized gauge
parameters could be set to zero, but a simpler and more insightful
result can be obtained through physical arguments.  $\mathcal{N}=4$
SYM provides several hints for further conditions to impose on the
ansatz.  For example, for maximally supersymmetric gauge theory it is
possible to enforce the no triangle hypothesis, manifest loop power
counting, and, for four-point and up to at least six loops, strip off
a factor of $st \atree$ from the integrand \cite{FiveLoopN4,
  Bern:2012uf, Bern:2010fy, SuperSum, Carrasco:2021otn,
  JJHenrikReview, BRY, BDDPR, Neq44np}.  For the NLSM integrand it is
desirable to make color-kinematics duality as manifest as possible.
One hope would be to factor out some piece of the one-loop numerator
since this at least manifests antisymmetry for the vertices involving
external legs.  However, a more fruitful direction is to match onto
the one known theory with pion power counting that manifests
color-kinematics duality to all loop orders, namely,
Zakharov-Mikhailov (ZM) theory \cite{Zakharov:1973pp, Cheung:2022mix}.
ZM theory is governed by the Lagrangian
\begin{equation}
\label{eq:ZMLagrangian}
\mathcal{L}^{\text{ZM}} = \frac{1}{2}(\partial \varphi)^2
+ g f^{abc} \varphi^a \varepsilon^{\mu\nu}(\partial_\mu \varphi^b)( \partial_\nu \varphi^c),
\end{equation}
where more details can be found in \cref{sec:CKLagrangians}.  The
color-stripped Feynman rule for the vertex is
$V(k_1, k_2, k_3) \propto \varepsilon_{\mu\nu}k_1^\mu k_2^\nu$ where
off-shell color-kinematics duality to all orders in perturbation
theory simply follows from the Schouten identity in 2D.  Since the theory
is purely cubic and manifests off-shell color-kinematics
duality, it is trivial to read off the color-dual numerator for any
graph.  From the presence of the Levi-Civita tensor
$\varepsilon^{\mu\nu}$, the theory clearly resides in two spacetime
dimensions where scattering is notoriously plagued by infrared
regulation issues.  Every on-shell particle is either a right mover,
with momentum proportional to $k_R^\mu \equiv (1,1)$, or a left mover,
with momentum proportional to $k_L^\mu \equiv (1,-1)$.  On-shell ZM
amplitudes naturally divide into sectors corresponding to the
configuration of left and right movers, where the scattering in many
sectors is rather subtle.  Only color-ordered amplitudes can be
defined unambiguously from the Feynman rules.  At four-point, only the
alternating sector is free of subtleties and the amplitude for this
process vanishes.  In equations, $A[LRLR]=0$ where $R$ corresponds to
a right mover and $L$ corresponds to a left mover.  In every other
configuration of right and left movers, such as $A[RRRR]$, one of the
internal propagators is accidentally on-shell since $k_R^2=k_L^2=0$.

>>>>

%\usepackage[notcite]{showkeys}
\usepackage{tikz-feynman} 
\usepackage{xcolor}
\hypersetup{
    colorlinks,
    linkcolor={nhpRed},
    citecolor={nhpBlue},
    urlcolor={nhpBlue}
}

@@

As noted above, we can select out can manifestly cubic
sector of semi-abelian Yang-Mills from the full theory of \cref{eq:ym-lag} by selecting only $(\varepsilon^+\varepsilon^+)$ in light-cone gauge (i.e. SDYM) and one-minus at tree-level or by plugging in the on-shell states of $J$-theory \cite{Cheung:2020djz,Cheung:2021zvb} and
$YZ$-theory \cite{Cheung:2016prv}. Indeed, semi-YM is just the following blocks
in the expansion of Yang-Mills given in Ref.~\cite{Pavao:2022kog}
\begin{equation}
A(A_1,...,\bar{A}_i,...,A_n)_{\text{tree}} = \sum_{i\neq j} (\varepsilon_i \varepsilon_j) \Delta^{(ij)}_{\text{tree}} \sim \sum_{i\neq j} (\varepsilon_i \varepsilon_j)  \sum_a \,[(\varepsilon k)^{n-2} (kk)^{3-n}]_a\,,
\end{equation}
and similarly so at one-loop
\begin{equation}
A(A_1,...,A_n)_{\text{1-loop}} = \Delta^{(\varnothing)}_{\text{1-loop}} \sim \sum_a\,[ (\varepsilon k)^{n} (kk)^{-n}]_a
\end{equation}
where $[\,\cdots]_a$ are terms with appropriate powers of $(\varepsilon k)$ and $(kk)$. \jm{Should we mention transmutation around here?} As we describe in detail in \cref{sec:CKLagrangians}, to recover NLSM at one-loop we need to extract the $D$-dependent part
that corresponds to an internal $\bar{Y}Y$-loop from extra-dimensional
scalars
\begin{eBox}
\begin{equation}
A^{\text{NLSM}}_{\text{1-loop}} \equiv \partial_D \Delta^{(\varnothing)}_{\text{1-loop}}\big|_{\varepsilon\rightarrow k}\, .
\end{equation}
\end{eBox}
Why must we take the derivative with respect to $D$? After all,
$J$-theory and $YZ$-theory both have well defined propagators for
internal $\bar{J}J$ and $\bar{Z}Z$ propagators. However, as we discuss
in \cref{sec:CKLagrangians}, all the $J$ and $Z$ states must be on-shell in order to
produce NLSM amplitudes. Thus, the unitarity cuts of $J$ theory at
one-loop will not produce NLSM amplitudes. However, the internal
$YY$-loop is a valid forward limit for producing NLSM amplitudes,
since as constructed $YZ$-theory matches to NLSM for off-shell
Y-particles. We now provide explicit expressions for these one-loop amplitudes in the next section.

\subsection{One-loop color-dual integrands}\label{oneLoopCK}
Our first application of this theory for color-dual construction at
loop level is for self-dual Yang-Mills (SDYM). As we have done throughout the text, we will set all coupling constants to unity. At tree-level, the
off-shell cubic vertices can be written in terms of light cone
coordinates \cite{Monteiro2011pc}
\begin{equation}
X(p,k) = p_u k_w-p_w k_u.
\end{equation}
At loop-level, this construction needs to be analytically continued to
general dimension in order to apply dimensional regularization at
one-loop. This can be done using the cubic semi-YM vertices of the
previous section
\begin{equation} \label{DdimSDYM}
\cubic{hgrey0}{fermion2}{}{fermion2}{}{fermion2}{}= \mathcal{X}(k_1,k_2) = (\varepsilon_1\bar{\varepsilon}_3)(\varepsilon_2 k_3) -  (\varepsilon_2\bar{\varepsilon}_3)(\varepsilon_1 k_3).
\end{equation}
Plugging in on-shell all-plus helicity states in light-cone gauge will
yield precisely the 4D SDYM vertex, up to an unphysical phase
  \begin{equation}
  \mathcal{X}(k^+_1,k^+_2) = \frac{\langle 12\rangle^3}{ \langle23\rangle \langle 31\rangle} \sim \langle 12\rangle = k_{1,u} k_{2,w}-k_{1,w} k_{2,u}\, ,
 \end{equation}
where $\mathcal{X}(k^+_1,k^+_2) = A^{\text{YM}}(1^+,2^+,3^-)$ in light cone gauge.
In the second equality, momentum conservation has been applied to redefine the spinor bracket, $\langle 12\rangle \rightarrow X(p,k)$, given in \cref{Conventions}.
\jm{Did I say this alright?}
Of course, the form of \cref{DdimSDYM} has the advantage of
permitting a $D$-dimensional construction of the one-loop
integrand. As an example, the four-point numerator is
\begin{equation}
\simplebox = \langle \mathcal{X}(k_1,\ell_1)\mathcal{X}(k_2,\ell_2)\mathcal{X}(k_3,\ell_3)\mathcal{X}(k_4,\ell_4)\rangle \, ,
\end{equation}
where $\ell_i = \ell-(k_1+k_2+\cdots+ k_i)$ and the bracket $\langle \,\cdots\rangle$ indicates that we have applied the gauge fixed state projector, $\sum {\varepsilon^\mu_{(+\ell_i)}\varepsilon^\nu_{(-\ell_i)}}=\eta^{\mu\nu}$, on all internal polarizations.
In general, the $n$-gon diagram is
\begin{equation}\label{SDYMnGon}
N^{\text{SDYM}}_{n\text{-gon}}= \nGonVector = \langle \mathcal{X}(k_1,\ell_1)\mathcal{X}(k_2,\ell_2)\cdots \mathcal{X}(k_n,\ell_n) \rangle\, .
\end{equation}
Recall that all other numerators can be obtained from the $n$-gon through Jacobi.
Due to the state-sum of internal loop factors,
$\sum {\varepsilon_{(+\ell)}\varepsilon_{(-\ell)}}\sim D$, the integrand above
depends explicitly on the spacetime dimension, $D$. By taking a
derivative\footnote{Another way to understand this derivative is that
  it selects the large $D$ behavior of the integrand.}, we can recover
the integrand needed for the all-plus one-loop amplitudes with an internal
scalar
\begin{equation}
\partial_D N^{\text{SDYM}}_{n\text{-gon}}=\nGonScalar  = 2^n (\ell_1 \varepsilon_1)(\ell_2 \varepsilon_2)\cdots (\ell_n \varepsilon_n)\, .
\end{equation}
This integrand numerator with internal scalar loop is precisely what
one would obtain from the Feynman rules of $YZ$-theory, absent the $\bar{Z}Z$
internal vector loop. For more background, we refer the reader to
\cref{sec:CKLagrangians}. When plugging on the on-shell states of $YZ$-theory, we thus obtain the following expression for the NLSM one-loop $n$-gon numerator
\begin{equation}\label{NLSMnGon}
N^{\text{NLSM}}_{n\text{-gon}}=\left[\partial_D  N^{\text{SDYM}}_{n\text{-gon}}\right]^{\epsilon \rightarrow k}=\nGonScalar =   [12][23]\cdots [n1]\, ,
\end{equation}
where we have define the antisymmetric kinematic variable,
$[ij] = \ell_i^2 - \ell_j^2 = 2(k_i \cdot \ell_i)$.
We have
verified through 10-point one-loop that this $n$-gon expression is a
valid color-dual representation for NLSM.

Before proceeding, we note that the above definition does give rise to ``pathological" bubble-on-external-leg (BEL) diagrams, discussed previously in \cref{sec:bootstrap}. However, one can show that these diagrams integrate to zero for spacetime dimension, $D>2$, and thus can be disregarded as unphysical. For the interested reader, in \cref{BELreg} we provide a detailed overview of this dimensional regularization of the relevant BEL diagram.
\subsection{Two-loop obstruction}\label{2LoopObstruction}
At two-loop, introducing terms that conspire with four-point contacts
is unavoidable. At one-loop, we were able to avoid internal
contractions of $\bar{A}_\mu A^\mu$ by selecting appropriate external
states. However, at two-loop when choosing all external $A_\mu$
states, the amplitude vanishes in semi-YM theory
\begin{equation}
\mathcal{A}^{\text{semi-YM}}_{\text{2-loop}}(A_\mu,A_\mu,A_\mu,A_\mu)=0,
\end{equation}
that is, the theory is one-loop exact.
In order to produce non-vanishing interactions, we would need to
reintroduce $D$-dimensional vertices from the diagonal group of
$G_L\times G_R$ that have the wrong weight. Doing so will necessarily permit internal
contractions of $\bar{A}_\mu A^\mu$,
\begin{equation}\label{2loopTension}
\mathcal{A}^{\text{YM}}_{\text{2-loop}}(A_\mu,A_\mu,A_\mu,A_\mu)=\doubleBoxObstructionA +\doubleBoxObstructionB+\cdots 
\end{equation}
where we used white dots to indicate interaction vertices of
weight\footnote{We choose the convention that the abelianized gauge
  fields carry weight $\mathcal{W}[\bar{A}] = -1$, and non-abelian
  vectors are $\mathcal{W}[\bar{A}] = +1$.}
$\mathcal{W}[\bar{A}\bar{A} A] = -1$, rather than the isolated semi-YM
vertex which always carries weight $\mathcal{W}[\bar{A}A A] =
+1$. This immediately runs into the difficulty of introducing the
four-point contact needed for color-kinematics to be satisfied on all
internal edges. By a simple weight counting argument, one can see that
including this wrong sign interactions is unavoidable at two-loop and
higher.

Thus, to construct two-loop numerators prescriptively, as we have done at one-loop, would require knowledge of
the full kinematic algebra for Yang-Mills off-shell.  Since this is presently unavailable, we will tackle the two-loop integrand using an ansatz approach.

\section{Two-loop four-point bootstrap}
\label{2loopBoot}
As we have just seen, it is possible to coerce tree numerators into
one-loop numerators at any multiplicity for pions and related
theories, but that these methods cannot be reapplied to generate
higher-loop numerators.  For pions in particular, the two-loop no-go statement
is only for \emph{one particular representation of the theory} so it does
not completely preclude the existence of a two-loop color-dual
integrand.  We thus turn to the color-dual bootstrap method
described in \cref{sec:bootstrap} to construct a color-dual
representation of two-loop four-point NLSM.  Because of the similarity
of the problem setup, we will also use the opportunity to revisit the
construction of a color-dual representation of pure YM, extending the
search space beyond what was covered in Ref.~\cite{Bern:2015ooa} to include the most general local ansatz.

Both NLSM and YM share the same cubic graph basis, and thus their
defining Jacobi relations lead to the same graph basis.
Ignoring tadpole and BEL graphs, there are 14 cubic four-point two-loop
graphs (see \cref{fig:MaxCuts}) which are related to each other via 21
Jacobi relations.  As mentioned in \cref{sec:bootstrap},
color-kinematics duality ensures that the numerator of every graph can
be expressed in terms of a basis of the double box and penta-triangle,
\begin{equation}
\label{eq:jacobi-basis}
\NLSMBasisDoubleBox \text{ and } \NLSMBasisPentaTriangle .
\end{equation}
Obviously, the actual numerator dressings and physical properties
of the two theories differ, so we discuss them separately below.

@@

For a scalar theory like NLSM the numerator only depends on dot
products of momenta.  The basis of momentum invariants for both of the
graphs can be taken to be
\begin{equation}
\label{eq:scalar-basis}
N^{\text{NLSM}}\left[
  \scalebox{0.75}{\NLSMBasisDoubleBox} \right]
\text{ and }
N^{\text{NLSM}}\left[\scalebox{0.75}{\NLSMBasisPentaTriangle} \right]\in
\left\{\begin{array}{cccc}
    k_{13},& k_{23},& k_{15},& k_{16}, \\
    k_{25},& k_{26},& k_{35},& k_{36}, \\
    k_{55},& k_{56},& k_{66}
  \end{array}
  \right\},
\end{equation}
where $k_{ij}$ means $k_i \cdot k_j$, $k_5=\ell_1$, and $k_6 = \ell_2$.
%%%%
% Basis for both diagrams
% {d[p[1], p[3]], d[p[1], p[6]], d[p[1], p[7]], d[p[2], p[3]],  d[p[2], p[6]], d[p[2], p[7]], d[p[3], p[6]], d[p[3], p[7]],  d[p[6], p[6]], d[p[6], p[7]], d[p[7], p[7]]}
%%%%
Every vertex in NLSM scales as $k^2$ so the numerator of each diagram
scales as $k^{12}$.  At this mass dimension there are 8,008
independent dot products of momenta.  With two basis graphs there is a
total of 16,016 ansatz parameters.\footnote{\texttt{Mathematica} is
  not suited to solving large sets of equations with sixteen thousand
  parameters, so a different method is desirable.  One possibility is
  to use a sprase solver, as often employed for integral reduction.
  However, these solvers typically assume that the matrix has a very
  low density.  While the equations in YM naturally separate into
  different ``helicity'' sectors (or $e_i \cdot e_j$ and
  $e_i \cdot k_j$ sectors in $D$ dimensions) which limits the
  cross-talk between equations and thus the effective density, there
  is no such separation of sectors in a scalar theory.  Thus the NLSM
  constraint equations were solved using a custom solver, working name
  \texttt{FiniteFieldSolve}, designed to exactly solve large linear
  systems of arbitrary density.  \texttt{FiniteFieldSolve} will be
  released publicly shortly.} Imposing the boundary Jacobi relations
eliminates all but 4,473 of the original free parameters.  Further
imposing graph symmetries reduces the number of free parameters to
1,243.

The final essential constraint is unitarity.  The two physical cuts
that need to be enforced, shown in \cref{fig:emu}, only involve
four-point vertices.  Notably this means that even at two loops,
unitarity is still only probing the on-shell four-point NLSM amplitude.
Unitarity cuts besides those shown in \cref{fig:emu} are either
pathological or vanish because they involve a three-point point on-shell
amplitude. In order to manifest the $\mathbb{Z}_2$
symmetry of NLSM at the integrand level, we also require that all
non-pathological cuts containing a three-point vertex (such as the maximal
cut) vanishe.  Imposing all relevant cuts leaves 765 free parameters
representing the generalized gauge freedom in the answer.

@@

At this point color-kinematics duality has been satisfied, so the
double copy will proceed without issue.  However, the plethora of
generalized gauge freedom parameters leaves the option for enforcing
additional aesthetic constraints.  All remaining generalized gauge
parameters could be set to zero, but a simpler and more insightful
result can be obtained through physical arguments.  $\mathcal{N}=4$
SYM provides several hints for further conditions to impose on the
ansatz.  For example, for maximally supersymmetric gauge theory it is
possible to enforce the no triangle hypothesis, manifest loop power
counting, and, for four-point and up to at least six loops, strip off
a factor of $st \atree$ from the integrand \cite{FiveLoopN4,
  Bern:2012uf, Bern:2010fy, SuperSum, Carrasco:2021otn,
  JJHenrikReview, BRY, BDDPR, Neq44np}.  For the NLSM integrand it is
desirable to make color-kinematics duality as manifest as possible.
One hope would be to factor out some piece of the one-loop numerator
since this at least manifests antisymmetry for the vertices involving
external legs.  However, a more fruitful direction is to match onto
the one known theory with pion power counting that manifests
color-kinematics duality to all loop orders, namely,
Zakharov-Mikhailov (ZM) theory \cite{Zakharov:1973pp, Cheung:2022mix}.
ZM theory is governed by the Lagrangian
\begin{equation}
\label{eq:ZMLagrangian}
\mathcal{L}^{\text{ZM}} = \frac{1}{2}(\partial \varphi)^2
+ g f^{abc} \varphi^a \varepsilon^{\mu\nu}(\partial_\mu \varphi^b)( \partial_\nu \varphi^c),
\end{equation}
where more details can be found in \cref{sec:CKLagrangians}.  The
color-stripped Feynman rule for the vertex is
$V(k_1, k_2, k_3) \propto \varepsilon_{\mu\nu}k_1^\mu k_2^\nu \equiv \langle k_1 k_2 \rangle$ where
off-shell color-kinematics duality to all orders in perturbation
theory simply follows from the Schouten identity in 2D.  Since the theory
is purely cubic and manifests off-shell color-kinematics
duality, it is trivial to read off the color-dual numerator for any
graph.  From the presence of the Levi-Civita tensor
$\varepsilon^{\mu\nu}$, the theory clearly resides in two spacetime
dimensions where scattering is notoriously plagued by infrared
regulation issues.  Every on-shell particle is either a right mover,
with momentum proportional to $k_R^\mu \equiv (1,1)$, or a left mover,
with momentum proportional to $k_L^\mu \equiv (1,-1)$.  On-shell ZM
amplitudes naturally divide into sectors corresponding to the
configuration of left and right movers, where the scattering in many
sectors is rather subtle.  Only color-ordered amplitudes can be
defined unambiguously from the naive Feynman rules.  At four-point, only the
alternating sector is free of subtleties and the amplitude for this
process vanishes.  In equations, $A[LRLR]=0$ where $R$ corresponds to
a right mover and $L$ corresponds to a left mover.  In every other
configuration of right and left movers, such as $A[RRRR]$, one of the
internal propagators is accidentally on-shell since $k_R^2=k_L^2=0$.
